Propositions as Sessions, Mechanically

### 2015-04-25 paper/presentation.tex
Fixed some issues with the frame title macro (more generic command).

\let\oldframetitle\frametitle
\renewcommand{\frametitle}[1]{
  \oldframetitle{#1}\vspace{-3.5cm}
}

@@

\begin{frame}
\oldframetitle{Session Types}
\begin{math}
  \begin{array}{cccc}
    S & \bnf  & \inpt{T}{S} & \mbox{receive value of type T, continue as S}\\
      & \altt & \outpt{T}{S} & \mbox{send value of type T, continue as S}\\
      & \altt & \Branch{l_i}{S_i}{i\in I} &
                           \mbox{offer between $|I|$ alternatives}\\
      & \altt & \Choice{l_i}{S_i}{i\in I} &
                           \mbox{choose between $|I|$ alternatives}\\
      & \altt & \mbox{\lstinline{end}} & \mbox{terminate session}\\
 \dual{S} & \bnf & \mbox{dual of S}\\
 \dual{(\inpt{T}{S})} & \bnf & \outpt{T}{\dual{S}}\\
 \dual{\Branch{l_i}{S_i}{i\in I}} & \bnf & \Choice{l_i}{\dual{S_i}}{i\in I}\\
            \vdots                & \cdots & \vdots
 \end{array}
\end{math}
\note{
one approach to achieve such correctness guarantees is by using binary session
types. [INTRODUCE BINARY SESSION TYPES HERE]. specifying a communication
protocol then involves providing mutual dual types to the communicating
parties.
}
\end{frame}

\begin{frame}[fragile]
\oldframetitle{GV Example}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} cc}
Process A ($A_x$) & Process B ($B_x$)\\
$\colored\tp{\inpt{Int}{\outpt{String}{\key{end}_!}}}$ ($\colored\tp{S}$) &
$\colored\tp{\outpt{Int}{\inpt{String}{\key{end}_?}}}$
  ($\colored\tp{\overline{S}}$)\\
\begin{gv}
let (m,$x_1$) = receive $x$ in
send thanks $x_1$
\end{gv} &
\begin{gv}
let $x_1$ = send gift $x$ in
let (resp,$x_2$) = receive $x_1$ in
terminate $x_2$
\end{gv}
\end{tabular*}
\[\colored
\inference{
  \Tp{thanks:\tp{String} \comma x:\tp{S}} \Tpvdash A_x:\tp{\key{end}_!} &
  \Tp{gift:\tp{Int} \comma x:\tp{\overline{S}}} \Tpvdash B_x:\tp{\key{Unit}}
}{Connect}{
  \Tp{thanks:\tp{String} \comma gift:\tp{Int}} \Tpvdash
    \Tm{with}~x~\Tm{connect}~A_x~\Tm{to}~B_x : \tp{Unit}
}
\]
\end{frame}

\begin{frame}
\frametitle{Motivation}
Hand-written (informal) proofs $=$ Error-prone $+$ Difficult to update
\note{
there has been a lot of activity in this area producing a variety of different
session-based type systems. many systems build on top of previous work adding
more complex features such as asynchronous, polymorphic and recursive session
types. extending informal proofs is error-prone and difficult because it is
not always clear which parts of a system's metatheory is affected by a
change. a formalised system would reduce likelihood of errors in the proofs
and aid extension since the theorem prover would highlight changes to the
metatheory
}
\end{frame}

\begin{frame}
\frametitle{System Overview}
CP $=$ process calculus with operators from classical linear logic

@@

GV $\to$ CP translation
\note{
the formalisation i present here is an encoding of a system in the Coq proof
assistant. it is based on a two-tier system providing a logical foundation for
understanding session types. A high-level overview of the system is as
follows: we have a process calculus CP with operators from classical linear
logic, a high-level functional language GV with session types, and a
translation from GV to CP which interprets operators in classical linear logic
as session types.
}
\end{frame}

@@

\begin{frame}
GV
\note{

>>>>

\let\oldframetitle\frametitle
\renewcommand{\frametitle}[1]{
  \oldframetitle{#1}\vspace{-3.5cm}
}
\newcommand{\ftitle}[2]{\oldframetitle{#1}\vspace{-#2cm}}

@@

\begin{frame}
\ftitle{Session Types}{1.5}
\begin{math}
  \begin{array}{cccc}
    S & \bnf  & \inpt{T}{S} & \mbox{receive value of type T, continue as S}\\
      & \altt & \outpt{T}{S} & \mbox{send value of type T, continue as S}\\
      & \altt & \Branch{l_i}{S_i}{i\in I} &
                           \mbox{offer between $|I|$ alternatives}\\
      & \altt & \Choice{l_i}{S_i}{i\in I} &
                           \mbox{choose between $|I|$ alternatives}\\
      & \altt & \mbox{\lstinline{end}} & \mbox{terminate session}\\
 \dual{S} & \bnf & \mbox{dual of S}\\
 \dual{(\inpt{T}{S})} & \bnf & \outpt{T}{\dual{S}}\\
 \dual{\Branch{l_i}{S_i}{i\in I}} & \bnf & \Choice{l_i}{\dual{S_i}}{i\in I}\\
            \vdots                & \cdots & \vdots
 \end{array}
\end{math}
\note{
one approach to achieve such correctness guarantees is by using binary session
types. [INTRODUCE BINARY SESSION TYPES HERE]. specifying a communication
protocol then involves providing mutual dual types to the communicating
parties.
}
\end{frame}

\begin{frame}
\ftitle{System Overview}{2}
CP $=$ process calculus with operators from classical linear logic

@@

GV $\to$ CP translation
\\~\\
Relies on Curry-Howard variant:
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}ccc}
Propositions & correspond to & Session Types\\
Proofs & correspond to & Processes\\
Cut Elimination & correponds to & Communication
\end{tabular*}
\note{
the formalisation i present here is an encoding of a system in the Coq proof
assistant. it is based on a two-tier system providing a logical foundation for
understanding session types. A high-level overview of the system is as
follows: we have a process calculus CP with operators from classical linear
logic, a high-level functional language GV with session types, and a
translation from GV to CP which interprets operators in classical linear logic
as session types.
}
\end{frame}

\begin{frame}[fragile]
\ftitle{GV Example}{2}
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}} cc}
Process A ($A_x$) & Process B ($B_x$)\\
$\colored\tp{\inpt{Int}{\outpt{String}{\key{end}_!}}}$ ($\colored\tp{S}$) &
$\colored\tp{\outpt{Int}{\inpt{String}{\key{end}_?}}}$
  ($\colored\tp{\overline{S}}$)\\
\begin{gv}
let (m,$x_1$) = receive $x$ in
send thanks $x_1$
\end{gv} &
\begin{gv}
let $x_1$ = send gift $x$ in
let (resp,$x_2$) = receive $x_1$ in
terminate $x_2$
\end{gv}
\end{tabular*}
\[\colored
\inference{
  \Tp{thanks:\tp{String} \comma x:\tp{S}} \Tpvdash A_x:\tp{\key{end}_!} &
  \Tp{gift:\tp{Int} \comma x:\tp{\overline{S}}} \Tpvdash B_x:\tp{\key{Unit}}
}{Connect}{
  \Tp{thanks:\tp{String} \comma gift:\tp{Int}} \Tpvdash
    \Tm{with}~x~\Tm{connect}~A_x~\Tm{to}~B_x : \tp{Unit}
}
\]
\end{frame}

\begin{frame}
\frametitle{Motivation}
Informal proofs $=$ Error-prone $+$ Difficult to update
\note{
there has been a lot of activity in this area producing a variety of different
session-based type systems. many systems build on top of previous work adding
more complex features such as asynchronous, polymorphic and recursive session
types. extending informal proofs is error-prone and difficult because it is
not always clear which parts of a system's metatheory is affected by a
change. a formalised system would reduce likelihood of errors in the proofs
and aid extension since the theorem prover would highlight changes to the
metatheory
}
\end{frame}

@@

\begin{frame}
GV is similar
\note{
